%!TEX root = thesis.tex

\chapter*{\vspace{-2.3cm} \Large Appendix C: Biography \vspace{1.7cm}}
\addcontentsline{toc}{chapter}{Appendix C: Biography}

\fancyhead[LO]{}
\fancyhead[RE]{Appendix C: Biography}

Bo{\v s}tjan Kalu{\v z}a received his Diploma in 2008 from the Faculty of Computer and Information Science, University of Ljubljana, Slovenia, by defending the thesis \textit{Analysis of Minimax Pathology and Pearl Game}. He was awarded with the Pre{\v s}eren Award (the highest faculty award) for his thesis.

In 2008, he enrolled in the \textit{New Media and E-Science} program at the Jo\v{z}ef Stefan International Postgraduate School, Ljubljana, Slovenia. He held a young researcher scholarship for doctoral studies and research training awarded by the Slovene Research Agency.

He spent an academic year as a visiting researcher at the University of Southern California in the Teamcore research group under the advisement of Prof.\ Dr.\ Milind Tambe, where he studied anomalous and suspicious agent behavior.

Since October 2008, he has been working at the Department of Intelligent Systems at the Jo\v{z}ef Stefan Institute, under the advisement of Prof. Dr. Matja\v{z} Gams. His research focuses on the development of novel algorithms and approaches, with an emphasis on %two domains: 
analysis of agent behavior from sensor data. For his work, he has received two best student paper awards, one at the International Conference on Machine Learning and Data Analysis (San Francisco, 2010), and the other at the 3rd Jo\v{z}ef Stefan International Postgraduate School Students' Conference (Ljubljana, 2011).
