
\chapter*{{\Large{\vspace{-2.3cm}Appendix C: Acknowledgments}}}

\noindent \fancyhead{}
\fancyfoot{} 
\fancyhead[RO]{\footnotesize \thepage}
\fancyhead[LO]{\footnotesize {Appendix C: Acknowledgments}}
\fancyhead[LE]{\footnotesize \thepage}
\fancyhead[RE]{\footnotesize {Appendix C: Acknowledgments}}
\addcontentsline{toc}{chapter}{Appendix C: Acknowledgments}

\noindent % First paragraph has no indentation.

\noindent First and foremost, I would like to thank my supervisor
Assoc. Prof. Dr. Peter Koro\v{s}ec and co-supervisor Assist. Prof.
Dr. Toma\v{z} Javornik for their guidance, understanding, and professional
assistance during these years. The insightful discussions and extensive
explorations have been essential for this work, which I could not
have carried out without their help. I am sincerely thankful for their
time spent in reviewing and editing my research work.

I wish to thank the members of the Computer Systems Department at
the Jo\v{z}ef Stefan Institute for providing me an enjoyable and
comfortable working environment. I am especially thankful to my office
mates Uro\v{s} Bole, Katerina Ta\v{s}kova and Vida Vuka\v{s}inovi\'{c}
for making my working time even more gratifying.

For their help, professional advice and countless testing data sets,
I am particularly thankful to the engineers of the Radio Network Department
at Telekom Slovenije, d.d. Without their support, much of my research
work would not have been possible.

I would like to express my gratefulness to Assoc. Prof. Dr. Tsuyoshi
Hamada for his thoughtful guidance, encouragement, and making my research
work at the Nagasaki University an unforgettable experience. In particular,
I would like to especially thank Dr. Felipe A. Cruz for his valuable
comments, suggestions and debates during our joint research work,
the results of which were presented in Chapter~\ref{chap:04-Framework-design-and-implementation}.
In this context, Assoc. Prof. Dr. Tsuyoshi Hamada acknowledges support
from the Japan Society for the Promotion of Science (JSPS) through
its Funding Program for World-leading Innovative R\&D on Science and
Technology (First Program). I would also like to thank the staff at
the Nagasaki Advanced Computing Center for their help and assistance
during my stay, and while making the computer cluster DEGIMA available
for the experimental simulations, the results of which provide a substantial
part of this thesis.

\bigskip{}


Last but not least, I wish to express my deepest gratitude and love
to my brother and my parents, for believing in me and always providing
their unconditional support.

\bigskip{}


All my research work, including this thesis, was co-financed by the
European Union, through the European Social Fund.
